Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems

### 2023-05-17 rmd/manuscript.Rmd
Update manuscript.Rmd

We employed a discrete-time model to describe tri-trophic dynamics in each habitat patch (see **Table 1** for parameter notations). A discrete model is suitable for organisms with predictable reproduction cycles, a common life history for many lotic species, including macroinvertebrates and fish. Our simulation model consists of three species; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**). The realized species abundance $A_{ix,t+1}$ for species $i$ (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:

@@

With no disturbance, the effects of ecosystem geometry were highly context dependent (**Figure 2**). In the chain scenario (no omnivory) with low productivity, ecosystem size had a positive effect on FCL (**Figure 3**), while ecosystem complexity had a negative influence (**Figure 4**). These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Figure S10**). However, when considering some levels of omnivory, these patterns characteristic of the chain scenario diminished. With weak omnivory (the second rows of **Figure 3 and 4**), ecosystem complexity had consistent positive effects on FCL, except the scenario with low productivity and short dispersal (**Figure 4**). In contrast, the effect of ecosystem size was more productivity dependent (**Figure 4**). Neither of ecosystem size and complexity influenced FCL with strong omnivory.

@@

<!-- ```{r} -->
<!-- source(here::here("code/table_param.R")) -->

<!-- knitr::kable(df_value %>% -->
<!--                dplyr::select(-"One-patch") %>%  -->
<!--                rename(Value = "Main"), -->
<!--              escape = FALSE, -->
<!--              booktabs = TRUE, -->
<!--              caption = "Parameter values used in the main simulation") %>%  -->
<!--   kableExtra::kable_styling(latex_options = 'HOLD_position') -->
<!-- ``` -->

>>>>

We employed a discrete-time model to describe tri-trophic dynamics in each habitat patch (see **Table 1** for parameter notations). A discrete model is suitable for organisms with predictable reproduction cycles, a common life history for many lotic species, including macroinvertebrates and fish. We assume non-overlapping generations of species because many lotic species, if not all, have semelparous life histories.

Our simulation model consists of three species; a basal species $B$ with positive logistic growth up to a carrying capacity, $K$; a primary consumer species or intraguild prey, $C$, which grows based on the consumption of $B$; and an intraguild predator $P$, which can consume both $B$ and $C$ (**Figure 1C**). The realized species abundance $A_{ix,t+1}$ for species $i$ (either $B,C,~or~P$) at patch $x$ and time $t+1$ is expressed as a composite of within-patch trophic interactions, immigration, and emigration:

@@

With no disturbance, the effects of ecosystem geometry were highly context dependent (**Figure 2**). In the chain scenario (no omnivory) with low productivity, ecosystem size had a positive effect on FCL (**Figure 3**), while ecosystem complexity had a negative influence (**Figure 4**). These opposing influences of ecosystem size and complexity were associated with their contrasting relationships with the total carrying capacity (i.e., carrying capacity summed across all patches): ecosystem size increased the total carrying capacity, whereas ecosystem complexity decreased it (**Figure S8**). However, when considering some levels of omnivory, these patterns characteristic of the chain scenario diminished. With weak omnivory (the second rows of **Figure 3 and 4**), ecosystem complexity had consistent positive effects on FCL, except the scenario with low productivity and short dispersal (**Figure 4**). In contrast, the effect of ecosystem size was more productivity dependent (**Figure 4**). Neither of ecosystem size and complexity influenced FCL with strong omnivory.

@@

```{r}
source(here::here("code/table_param.R"))

knitr::kable(df_value %>%
               dplyr::select(-"One-patch") %>%
               rename(Value = "Main"),
             escape = FALSE,
             booktabs = TRUE,
             caption = "Parameter values used in the main simulation") %>%
  kableExtra::kable_styling(latex_options = 'HOLD_position')
```
